Upper Ouémé Forest Fungi -- Soil Transect Analysis

### 2019-02-11 R/transect-analysis.Rmd
Restructure to fit standard Snakemake project structure with ./scripts and ./config directories, part 1

```{r}
# Load the IonTorrent OTU table
ion.otu <- read_tsv(iontorrent.otu.file) %>%
  mutate(n_reads = rowSums(select_if(., is_numeric))) %>%
  left_join(read_tsv(iontorrent.classification.file)) %>%
  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
  left_join(read_tsv(iontorrent.funguild.file) %>% rename(OTU = Group)) %>%
  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
  rename(Kingdom = Domain) %>%
  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "?") %>%
  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "unidentified") %>%

  # make "unident." as informative as possible
  dplyr::mutate_at("Kingdom", replace_na, "unident.") %>%
  dplyr::mutate(Phylum = ifelse(is.na(Phylum),
                                ifelse(is.na(Kingdom),
                                       NA_character_,
                                       paste0("unident. ", Kingdom)),
                                Phylum),
                Class = ifelse(is.na(Class),
                                ifelse(is.na(Phylum),
                                       NA_character_,
                                       paste0("unident. ", Phylum)),
                                Class),
                Order = ifelse(is.na(Order),
                                ifelse(is.na(Class),
                                       NA_character_,
                                       paste0("unident. ", Class)),
                                Order),
                Family = ifelse(is.na(Family),
                                ifelse(is.na(Order),
                                       NA_character_,
                                       paste0("unident. ", Order)),
                                Family),
                Genus = ifelse(is.na(Genus),
                                ifelse(is.na(Family),
                                       NA_character_,
                                       paste0("unident. ", Family)),
                                Genus),
                Species = ifelse(is.na(Species),
                                ifelse(is.na(Genus),
                                       NA_character_,
                                       paste0(Genus, " sp.")),
                                Species)) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "(unident\\. )*unident(ified|\\.)",
            "ZZZZunident.") %>%

  # Sort taxonomically. Put the "unident." at the end
  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "ZZZZ",
            "") %>%

  # Add the representative sequence for each OTU and calculate the length
  left_join(
    seqinr::read.fasta(file = iontorrent.repseq.file, as.string = TRUE) %>%
    { tibble(OTU = names(.),
             RepSeq = unlist(.),
             SeqLen = str_length(RepSeq)) }) %>%

  # Remove the positive control
  filter(is.na(Species) | Species != "Agaricus bisporus") %>%

  # Keep the data in this order.
  mutate_at(dplyr::vars(Kingdom:Species),
            ~ factor(.) %>% fct_inorder)

ion.platemap <- read_tsv(iontorrent.otu.file) %>%
  column_to_rownames("OTU") %>%
  as.matrix %>%
  colSums %>%
  tibble(Well = names(.), n_reads = .) %>%
  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE) %>%
  full_join(plate %>%
              select(Plate, Well, sample_type) %>%
              full_join(
                expand.grid(Plate = 1:2,
                            Row = LETTERS[1:8],
                            Column = 1:12) %>%
                  tidyr::unite(Well, Row, Column, sep = ""))) %>%
  mutate(Plate = "IonTorrent") %>%
  mutate_at("sample_type", fct_explicit_na, "None") %>%
  group_by_if(~ !is.factor(.)) %>%
  summarize(sample_type = paste(as.character(sort(unique(sample_type))), collapse = "/")) %>%
  ungroup %>%
  mutate(f_reads = n_reads / max(n_reads)) %>%
  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
  mutate_at("Row", factor, levels = LETTERS) %>%
  mutate_at(c("Column", "Row"), as.integer)
```

@@

pb.otumatrix <- pb.otu %>%
  column_to_rownames("Group") %>%
  select(A7_483_1:H7_483_1) %>%
  as.matrix %>%
  t %>%
  as.tibble(rownames = "Sample") %>%
  tidyr::extract("Sample", c("Well", "Plate"), regex = "([A-H]1?[0-9])_483_([12])", remove = FALSE) %>%
  mutate_at("Plate", as.integer) %>%
  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE)

>>>>

@@

pb.otumatrix <- readRDS(file.path(data.dir, "short-pacbio_pb_483.dada.seqtable.rds"))

  pb.otu %>%
  column_to_rownames("Group") %>%
  select(A7_483_1:H7_483_1) %>%
  as.matrix %>%
  t %>%
  as.tibble(rownames = "Sample") %>%
  tidyr::extract("Sample", c("Well", "Plate"), regex = "([A-H]1?[0-9])_483_([12])", remove = FALSE) %>%
  mutate_at("Plate", as.integer) %>%
  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE)
